---
title: "16s_MS_steps"
author: "Shihan Zhang"
date: '2022-08-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Build virtul environment for new project
```{r}
renv::init()
#renv::restore()
```

# Load packages
```{r}
library(readxl)
library(phyloseq)
library(decontam)
library(metagenomeSeq)
library(vegan)
#library(ANCOMBC) need fix
library(ggplot2)
library(ggside)
library(ggsignif)
library(ggpubr)
library(gridExtra)
library(fantaxtic)
library(dplyr)
library(tidyverse)
library(hydroTSM)
library(lubridate, warn.conflicts = FALSE)
#library(DESeq2)
library(edgeR)
library(Maaslin2)

# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("decontam")

# save package setups
# renv::snapshot()
```

### ggplot setting
```{r}
# color code
custom_palette <- c("cadetblue4","lightsalmon3","dodgerblue2",'gold','deeppink','red4','purple','green4','black')
  # color setting
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector[13] <- 'firebrick'
col_vector[19] <- 'wheat3'
col_vector[20] <- 'cyan3'
```


# I6S Data Import
```{r data import}
# Load data from ITS pipeline outputs
# sequence table - OTU table - OTU counts for each sample
bact.seqtab <- t(data.frame(readRDS('/Users/senmnes/DESKTOP_HD/Bernet_6th/Multi_Omics/raw_16S/seqtab.rds')))

# taxonomy table
bact.taxonomy <- data.frame(readRDS('/Users/senmnes/DESKTOP_HD/Bernet_6th/Multi_Omics/raw_16S/taxonomy_species.rds'))

# phylogenetic tree file
bact.tree <- readRDS('/Users/senmnes/DESKTOP_HD/Bernet_6th/Multi_Omics/raw_16S/tree.rds')

# clinical metadata
metadata <- read_csv('/Users/senmnes/DESKTOP_HD/Bernet_6th/MS_downstream/Metadata_2022.csv')

# Sample Status data
samplestatus <- read_csv('/Users/senmnes/DESKTOP_HD/Bernet_6th/Multi_Omics/Sample_status - 16S.csv')
```


## Data Wrangling
Minor adjustment for original observations for easier ploting in later steps
```{r}
# Add groups
# Rename clip identification
metadata$WheezeClip <- ifelse(metadata$WheezeClip1YEAR=='ConfirmedWheeze', 'Yes right clip',ifelse(metadata$WheezeClip1YEAR=='WrongClipWheeze', 'Other clip', 'No clip'))
# Add a 'Wheeze' column merging confirmed and unconfirmed wheezers
metadata$Wheeze <- ifelse(metadata$WheezeClip1YEAR=='No', 'No', 'Yes')
# add groups
metadata$Group <- ifelse(metadata$Wheeze =='No', 'Healthy', 'Wheeze')

# Clinical covariates relevant for 1 year time point
clinical_factors <- c('WheezeClip','DoctorVisitsAsthmaWheeze1YEAR','EmergencyVisitsAsthmaWheeze1YEAR','HospitalAdmissionsAsthmaWheeze1YEAR','DiagnosedBronchiolitis1YEAR','ChestInfections1YEAR','Colds1YEAR','DryCoughAtNight1YEAR','Sneezing1YEAR','FoodAdverseReactions1YEAR','Rash1YEAR')

# Genetic/hereditary covariates relevant for all time points
genetic_factors <- c('MotherAsthma','FatherAsthma','MotherEczema','FatherEczema',
                     'MotherHayfever','FatherHayfever',
                     'Gender')

# Environmental covariates relevant for 1 year time point
environmental_factors <-c('DayCare1YEAR','TimeOfYear','NbHouse','ParentsSmoking','PetsAtHome1YEAR','BreastfeedingBirth','AbxSinceBorn1YEAR','Country')

# Birth/pregnancy covariates relevant for all time points (particularly early ones)
birth_pregnancy_factors <-c('DeliverySimplified','GestAgeBirth','WeightBirth','AbxCoursesPregnancy','BreastfeedingBirth')

# adjustments 
metadata$PetsAtHome1YEAR <- ifelse(metadata$PetsAtHome1YEAR == 'None', 'No', 'Yes')
metadata$ParentsSmoking <- ifelse(metadata$FatherSmoke != 'Noneversmoked'|metadata$MotherSmoke != 'Noneversmoked', 'Yes', 'No')
metadata$EthnicitySimplified <- ifelse(metadata$Ethnicity == 'WhiteBritish','British','Other')

# Antibiotics Courses history
#metadata$AbxSinceBorn1YEAR <- ifelse(metadata$AbxSinceBorn1YEAR == 0, 'No', 'Yes')
#metadata$AbxSinceLastSeen3MTH <- ifelse(metadata$AbxSinceLastSeen3MTH == 'No', 'No', 'Yes')
#metadata$AbxSinceLastSeen6MTH <- ifelse(metadata$AbxSinceLastSeen6MTH == 'No', 'No', 'Yes')
```

### Adding Age & Season
```{r}
# add season
date_factor <- c('DateofVisitBaseline','DateofVisit1YEAR','DateofVisitAcute','DateofVisit3MTH','DateofVisit6MTH')

# reformat date
for(date_col in date_factor){
  #print(date_col)
  metadata[[date_col]] <- ymd(metadata[[date_col]])
}
# adjust date formate
metadata$ParticipantsDateofbirth <- ymd(metadata$ParticipantsDateofbirth)
metadata$month1YEAR <- month(metadata$DateofVisit1YEAR)
metadata$TimeOfYear <- ifelse(metadata$month1YEAR>=5 & metadata$month1YEAR<=10, 'Warm', 'Cold')

# calculate age in month
# 0
metadata$AgeofVisitBaseline <- day(as.period(interval(start=metadata$ParticipantsDateofbirth,end=metadata$DateofVisitBaseline), unit='days'))/30.437
# 3
metadata$AgeofVisit3MTH <- day(as.period(interval(start=metadata$ParticipantsDateofbirth,end=metadata$DateofVisit3MTH), unit='days'))/30.437
# 6 
metadata$AgeofVisit6MTH <- day(as.period(interval(start=metadata$ParticipantsDateofbirth,end=metadata$DateofVisit6MTH), unit='days'))/30.437
# 1YEAR
metadata$AgeofVisit1YEAR <- day(as.period(interval(start=metadata$ParticipantsDateofbirth,end=metadata$DateofVisit1YEAR), unit='days'))/30.437
# ACUTE
metadata$AgeofVisitACUTE <- day(as.period(interval(start=metadata$ParticipantsDateofbirth,end=metadata$DateofVisitAcute), unit='days'))/30.437

# At 1YEAR age, 1YEAR is the last observation collection point
metadata$At1YEARpoint_Age <- NA
for(i in seq_along(metadata$AgeofVisit1YEAR)){
  if(is.na(metadata$AgeofVisit1YEAR[i])==F){
      if(between(metadata$AgeofVisit1YEAR[i],0,3)){
        metadata$At1YEARpoint_Age[i] <- '<3'
      }else if(between(metadata$AgeofVisit1YEAR[i],3,6)){
        metadata$At1YEARpoint_Agep[i] <- '3-6'
      }else if(between(metadata$AgeofVisit1YEAR[i],6,12)){
        metadata$At1YEARpoint_Age[i] <- '6-12'
      }else if(between(metadata$AgeofVisit1YEAR[i],12,20)){
        metadata$At1YEARpoint_Age[i] <- '>12'
      }
  }
}

# convert to season
temp_date <- metadata$DateofVisit1YEAR
metadata$Season1YEAR  <- str_to_title(time2season(temp_date,out.fmt="seasons"))
# fix typo
for(rows in seq_along(metadata$Season1YEAR)){
  if (is.na(metadata$Season1YEAR[rows])){
  }else if(metadata$Season1YEAR[rows] == 'Autumm'){metadata$Season1YEAR[rows] = 'Autumn'}
}
```

## Extract Target barcodes
Build up Sample Status barcode df
- 'Baseline' all green
- 'Baseline-NP' - all red
- 'xxxxxxx-BLUE' - BLUE barcodes
```{r}
# Add 16S tag for all sample status barcode
for (cols in colnames(samplestatus)[-1]){
  samplestatus[[cols]]<-ifelse((nchar(samplestatus[[cols]]) <= 7 | is.na(samplestatus[[cols]]) == T),NA,samplestatus[[cols]] <- paste('16S',sep='.',samplestatus[[cols]]))
}
# long format is the correct way to assign groups
# Transform to long format
long_ss <- samplestatus %>% 
  pivot_longer(
    cols = Baseline:`Acute-NP`,
    names_to='SampleCollection',
    values_to='Barcode')

# make a clean SS df without any BLUEs.
long_ss <- long_ss[(grepl('BLUE',long_ss$Barcode)==F),]
long_ss <- long_ss[!is.na(long_ss$Barcode),]
```

# Before PhyloSeq
## Metadata transformation - before any wrangling steps to prevent index miss match
This part requires `library(tidyverse)`
1. Added negative control/barcodes into metadata. If the barcode IN the seqtab but NOT in the metadata, wee add these barcodes back into metadata.
 
2. assign barcodes to be rowname for input for phyloseq package.
- Reorder the barcodes. The barcodes from Metadata should follows the exactly same order as the Seqtab. 

2.1 Label Negative controls. After adding the barcodes back to metadata, we now need to label the negative control with an extra column `IsNegativeCtrl`  
 - Only `PCR` and `ExtrNeg` are negative controls. !!! Not include `Zymo`
 - some Outliers observation like 'Notdone'/'Nitdone'/'refused' are originally from the data collection.
 
- In this step, we will facing a problems which is the duplicated barcodes. In R, we are not allowed to have duplicated rownames. But at this stage, we are not planned to remove any duplicates barcodes. Therefore, we will back up the duplicated barcodes first, then remove the duplicated barcodes(sample) and NAs.

```{r pre-process metadata}
# Step 1
# Transform the metadata with extract barcode as key
barcodeList<-c(grep('BCNasalswab', colnames(metadata), value=TRUE))
bcd_all <- c(grep('BCNasalbrushing', colnames(metadata), value=TRUE))
# remove the barcode not for this study
for(i in bcd_all){
  metadata[[i]] <- NULL
}

# Transform to long format
metadata <- metadata %>% 
  pivot_longer(
    cols = all_of(barcodeList), 
    names_to='SampleCollection',
    values_to='Barcode')
```

## Matching 16S.Seqtab with Sample Status file
The barcodes not in the sample status file are not relevant with this study, we dont want them.
But Seqtab contains negative control, we want to keep negative control but remove other barcode that not exsit in SS.
```{r}
# find the barcodes in ss but not in seqtab
n_seq_y_ss <- colnames(bact.seqtab)[!is.element(long_ss$Barcode,colnames(bact.seqtab))]
# in seqtab but not in ss
y_seq_n_ss <- colnames(bact.seqtab)[!is.element(colnames(bact.seqtab),long_ss$Barcode)]
# These barcodes will contain negative controls, we want to keep negative controls
for(i in seq_along(n_seq_y_ss)){
  if(grepl('PCRCtrl', n_seq_y_ss[i])){n_seq_y_ss[i] <- 0}
  else if(grepl('ExtrNegCtrl', n_seq_y_ss[i])){n_seq_y_ss[i] <- 0}
  else if(grepl('ZymoCtrl', n_seq_y_ss[i])){n_seq_y_ss[i] <- 0}
  else{}
}

for(i in seq_along(y_seq_n_ss)){
  if(grepl('PCRCtrl', y_seq_n_ss[i])){y_seq_n_ss[i] <- 0}
  else if(grepl('ExtrNegCtrl', y_seq_n_ss[i])){y_seq_n_ss[i] <- 0}
  else if(grepl('ZymoCtrl', y_seq_n_ss[i])){y_seq_n_ss[i] <- 0}
  else if(grepl('NegCtrlBAL', y_seq_n_ss[i])){y_seq_n_ss[i] <- 0}
  else{}
}

# clean up negative control and NAs
n_seq_y_ss <- n_seq_y_ss[(n_seq_y_ss!='0')]
n_seq_y_ss <- n_seq_y_ss[!is.na(n_seq_y_ss)]
y_seq_n_ss <- y_seq_n_ss[(y_seq_n_ss!='0')]
y_seq_n_ss <- y_seq_n_ss[!is.na(y_seq_n_ss)]

# filter bact.seqtab, remove the barcodes thats not in the ss
bact.seqtab <- bact.seqtab[,!(colnames(bact.seqtab) %in% y_seq_n_ss)]
dim(bact.seqtab) # raw seq.tab has (10814,441), now has (10814,409)
# check again, only return negative control group
colnames(bact.seqtab)[!is.element(colnames(bact.seqtab),long_ss$Barcode)]
```
Here we have very interesting result, There are 2 type of matching we have done. 
`n_seq_y_ss` has the barcode that appear in the SS but not in seqtab
`y_seq_n_ss` has the barcode that appear in the Seqtab but not in SS

- Given we trust in SS rather than Seqtab, so we can not add `y_seq_n_ss` back into SS.
What we may do is remove the `y_seq_n_ss` so Seqtab can be more clear.
- On the otherhand, `n_seq_y_ss` shows the barcode not having OUTs in the seqtab, because they not exsist in seqtab.

---Metadata Transformation - Continues---
This part requires `library(tidyverse)`
1. Added negative control/barcodes into metadata. If the barcode IN the seqtab but NOT in the metadata, wee add these barcodes back into metadata.
- some Outliers observation like 'Notdone'/'Nitdone'/'refused' are originally from the data collection. We want to remove them as well as the NAs.

- 1.1. Filter the metadata by `Sample_Status_Wheeze` file
  - We trust `Sample_Status_Wheeze` contains all the available samples, So we use SS to filter the metadata & Seqtab[done in above].
  - But need to clear is, We filter metadata by SS first, add negative control back as second. Because SS DOSE NOT contains negative controls. 

2. Label Negative controls. After adding the barcodes back to metadata, we now need to label the negative control with an extra column `IsNegativeCtrl`  
 - Only `PCR` and `ExtrNeg` are negative controls. !!! Not include `Zymo`
 
```{r}
# Step 1
# Extract the barcodes from seqtab
seqtab_barcodes <- colnames(bact.seqtab)

# Add back the tag '16S', without editing NA and also assign the outliers to NA
metadata$Barcode <- ifelse((nchar(metadata$Barcode) <= 7 | is.na(metadata$Barcode) == T),NA,metadata$Barcode <- paste('16S',sep='.',metadata$Barcode))

# Check the barcodes in Seqtab but not in metadata, Should only return Negative controls
missing <- seqtab_barcodes[!(is.element(seqtab_barcodes, metadata$Barcode))]

# Filter metadata base on the SS, PLEASE NOTE, metadata CAN NOT filter by bact.seqtab as the bact.seqtab contains less available barcodes than SS.
dim(metadata)
metadata <- metadata[is.element(metadata$Barcode,long_ss$Barcode),]
dim(metadata)

# Add Negative controls them back into metadata
metadata <- metadata %>% add_row(Barcode = c(missing))
dim(metadata)

# Double check with SS
metadata[!is.element(metadata$Barcode,long_ss$Barcode),]$Barcode # should only return negtaive controls

# Step 2
# label the negative controls with extra column`IsNegativeCtrl`
metadata$IsNegativeCtrl <- ifelse((grepl('PCRCtrl', metadata$Barcode)|grepl('ExtrNegCtrl', metadata$Barcode)), 'YES','NO')
metadata$IsNegativeCtrl <- ifelse(grepl('NegCtrlBAL', metadata$Barcode), 'YES','NO')

# Step 2.1 reorder Dataframe[metadata] by the order of barcode of bact.seqtab
#metadata <- metadata[match(colnames(bact.seqtab),metadata$Barcode),]
seq_name <- colnames(bact.seqtab)
# merge, but is has duplicated value,so use unique()
new_order <- unique(c(seq_name,metadata$Barcode))
# now, the `new_order` following the extact order of bact.seqtab with no removed barcode by match()
metadata <- metadata[match(new_order,metadata$Barcode),]

# re-assign the metadata to be a dataframe format.
# Even the matadata looks like a dataframe in R, when we import the data, we using RDS fromats. So, right here we will re assign the metadta to be a dataframe file to prevent any bugs for phyloseq object generation. THIS IS REALLY IMPORTANT!!!
metadata <- as.data.frame(metadata)
# assign barcode to be rowname of metadata
rownames(metadata) <- metadata$Barcode
```

#### Sample Status Data checking
```{r}
long_ss$Group <- NA
# assign group values into long_ss
for(i in long_ss$Barcode){
  ss_barcd <- i
  # get the match first, will return index
  idx <- match(ss_barcd,metadata$Barcode)[1]
  # by index, get value of group
  grp_info <- metadata$Group[idx][1]
  #  
  idx_ss <- match(ss_barcd,long_ss$Barcode)[1]
  long_ss$Group[idx_ss] <- grp_info
}
  
# manually fix these 2 data; they are duplicated barcodes
long_ss[(long_ss$Barcode=='16S.IBG102152A'),]$Group <- 'Wheeze'
long_ss[(long_ss$Barcode=='16S.EDG101702A'),]$Group <- 'Healthy'

# no diagnosis group
no_diagnosis <- long_ss[is.na(long_ss$Group),]

# # wheeze
ss_wheeze <- long_ss[(long_ss$Group == 'Wheeze'),]
# # healthy
ss_healthy <- long_ss[(long_ss$Group == 'Healthy'),]
```

## Function to assign taxonomy id
This function allows the creation of a new taxon name (ID) containing the highest taxonomic level to which the ASV has been assigned to.
```{r}
highest_taxClass <- function (x) {
  y <- ifelse(is.na(x$Species), ifelse(is.na(x$Genus),
                ifelse(is.na(x$Family),
                       ifelse(is.na(x$Order),
                              ifelse(is.na(x$Class),
                                     ifelse(is.na(x$Phylum),
                                            paste0("k.", x$Kingdom, "_", c(1:dim(x)[1])),
                                          paste0("p.", x$Phylum, "_", c(1:dim(x)[1])) ),
                                     paste0("c.", x$Class, "_", c(1:dim(x)[1])) ),
                                  paste0("o.", x$Order, "_", c(1:dim(x)[1])) ),
                       paste0("f.", x$Family, "_", c(1:dim(x)[1])) ),
                paste0("g.", x$Genus, "_", c(1:dim(x)[1])) ),
         paste0("s.", x$Genus, "_", x$Species, "_", c(1:dim(x)[1])) )
  return(y)
}
bact.taxonomy$ID <- highest_taxClass(bact.taxonomy)
```

# Build PhyloSeq
## Combine data into phyloseq object for downstream analysis
This step will combine all the data into phyloseq object. The object `bact` will carry have a comprehensive information.
```{r}
bact <- phyloseq(
          otu_table(bact.seqtab, taxa_are_rows = TRUE),
          sample_data(metadata),
          tax_table(as.matrix(bact.taxonomy)),
          phy_tree(bact.tree$tree)
          )

# rename the ASVs by taxa ID(highest taxonomic class)
taxa_names(bact) <- tax_table(bact)[,'ID']
```

# Plot relative abundance
```{r}
PRA <- function(plsq_obj,plt_name){
  # subset 1YEAR
  plsq_RA <- subset_samples(plsq_obj, SampleCollection == "BCNasalswab1YEAR")
  # select genus level
  plsq_genus <- tax_glom(plsq_RA,"Genus")
  # get top 20
  top <- names(sort(taxa_sums(plsq_genus), decreasing = TRUE))[1:20]
  # Calculate relative abundance
  #Bact_genus.prop <- transform_sample_counts(Bact_genus, function(x) x / sum(x) )
  plsq_genus.prop <- microbiome::transform(plsq_genus, "compositional")
  # filter the top 20
  plsq_genus.prop.20 <- prune_taxa(top, plsq_genus.prop)
  
  # Remove the "s.g__","g.g__" before the genus names.
  taxa_names(plsq_genus.prop.20) <- gsub("g.", "", taxa_names(plsq_genus.prop.20))
  taxa_names(plsq_genus.prop.20) <- gsub("s.", "", taxa_names(plsq_genus.prop.20))
  # Remove the number behind
  #taxa_names(Bact_genus.prop.20)<-gsub("[[:print:]][[:digit:]]","",taxa_names(Bact_genus.prop.20))
  #taxa_names(Bact_genus.prop.20)<-gsub("[[:digit:]]","",taxa_names(Bact_genus.prop.20))
  
  plot_composition(plsq_genus.prop.20, sample.sort = NULL,
                            otu.sort = NULL,
                            plot.type = "barplot",
                            verbose = FALSE) +
    labs(x='',y="Relative abundance",fill='Genus', title = paste(plt_name,'Top 20 taxa')) +
    theme_light() + 
    theme(axis.text.x = element_blank(),
          axis.ticks.x=element_blank(),
          legend.text = element_text(size=6),
          legend.key.size = unit(4, 'mm')) +
    scale_fill_manual(values = col_vector) + 
    theme(legend.position = 'right')
}
```

```{r}
bact.BF <- bact
#
pdf(file=paste0('/Users/senmnes/DESKTOP_HD/Bernet_6th/pre-process_plot','/BF.Bact1Y_top20.pdf'), width=8, height=4, useDingbats=F, compress=F)
#
  PRA(bact.BF,'Bacteria')
#
dev.off()
```

# After Build PhyloSeq
## Pre-processing 
This step will remove very low read samples. Also, the samples are decontaminated based on the negative controls composition using `decontam R package`

1). Sample filtering based on the read counts(first round). Samples with less than 100 reads(even negative controls) need to be removed. As chances are hight that they just represent sequecing background. Not removing them can impact the read decontamination process.
```{r}
# check samples number
print(paste('before FIRST ROUND removal: ',nsamples(bact), 'samples'))

# PLEASE NOTE!, the threshold number isnt same for all the cases, we need to manually test and found the capable one. 
minreadsTreshold <- 200

# filtering based on minimum threshold, colSums return a total reads count for each sample.
bact <- prune_samples(colSums(otu_table(bact)) > minreadsTreshold, bact)

# check samples number
print(paste('after  FIRST ROUND removal: ',nsamples(bact), 'samples'))
```

2). Remove unclassified ASVs. If the ASV is not assigned at Phylum level(NA), it is high chance that this seqence is not a bact(e.g. host-derived). These ASVs will be removed.
```{r}
# check the total counts of taxa
print(paste('before Unclassified ASVs removal: ',ntaxa(bact), 'taxonomy'))

# only keep the ASVs which the Phylum is not NA.
bact <- subset_taxa(bact, (!is.na(Phylum)))

# check the total counts of taxa
print(paste('after  Unclassified ASVs removal: ',ntaxa(bact), 'taxonomy'))
```

3). ASVs decontamination 
For microbiome data pre-processing(especially when working with low biomass samples). Decontamination is the key step! The `decontam` R package is used. with the prevalence method,
as no absolute quantification is available. The threshold can be adjusted depending on the dataset (usually between 0.1 and 0.5 - less strict to more strict).
- The function requires a vector (is.neg) to be TRUE to distinguish negative controls from the rest of the samples.
```{r}
# decontamination treshold
threshold <- 0.5
# negative control determine vector with pure Boolean value
sample_data(bact)$is.neg <- ifelse(grepl('PCRCtrl',sample_data(bact)$Barcode)|
                                    grepl('ExtrNegCtrl',sample_data(bact)$Barcode), TRUE,FALSE)
# distinguish negative controls
contamdf.prev <- isContaminant(bact, method='prevalence', neg='is.neg',threshold=threshold)
# assign contam samples
contam <- contamdf.prev[which(contamdf.prev$contaminant==TRUE),]

# check how many contaminating ASVs
table(contamdf.prev$contaminant)
```

Among possible plots, one can represent the prevalence of ASVs in negative controls vs actual samples and color code based on whether the ASV has been identified as a contaminant or not.
```{r}
# Make a new Phyloseq object of presence-absence in negative controls and true samples
bact_pa <- transform_sample_counts(bact, function(abund) 1 * (abund > 0))
bact_pa_neg <- prune_samples(sample_data(bact_pa)$is.neg == TRUE, bact_pa)
bact_pa_pos <- prune_samples(sample_data(bact_pa)$is.neg == FALSE, bact_pa)

# Make a data.frame of prevalence in positive and negative samples
df_bact_pa <- data.frame(bact_pa_pos=taxa_sums(bact_pa_pos),
                        bact_pa_neg=taxa_sums(bact_pa_neg),
                        contaminant=contamdf.prev$contaminant)

# Plot the prevalence of taxa between positive and negative samples
png(file=paste0('/Users/senmnes/DESKTOP_HD/Bernet_6th/pre-process_plot','/Bact_containminant.png'), width=1410, height=1060, unit='px')

ggplot(df_bact_pa, aes(x = bact_pa_neg,y = bact_pa_pos, color=contaminant)) +
    geom_point(size = 7) +
    labs(title = 'Bacteria decontam',
    x = 'Prevalence (Negative Controls)',
    y = 'Prevalence (True Samples)') +
    scale_color_manual(name = 'Contaminant', values = c('darkseagreen', 'indianred'))+
    theme_bw() + 
    theme(text = element_text(size=30))+
    theme(legend.position = 'right',
          legend.background = element_rect(linetype="solid",colour ="black"),
          legend.key.size = unit(1, "cm")) + guides(alpha="none")

dev.off()
```

## ASVs decontamination - con.
Remove ASVs decontamination - contaminants ASVs:
```{r}
# Remove the contaminants ASVs:
# check taxa counts
print(paste('before Contaminants ASVs removal: ',ntaxa(bact), 'taxonomy'))

contam_taxa <- c(rownames(contam))
allTaxa <- taxa_names(bact)
allTaxa <- allTaxa[!(allTaxa %in% contam_taxa)]
bact <- prune_taxa(allTaxa,bact)
# check taxa counts
print(paste('after  Contaminants ASVs removal: ',ntaxa(bact), 'taxonomy'))
```

5) ASVs filtering based on Prevalence filtering allows the removal of rare ASVs. This steps can greatly improve downstream analysis such as differential abundance testing and multi-omics data inte-gration. mincountsThreshold represents the minimum ASV count and prevalenceTheshold the minimum proportion of samples that should have the minimum ASV count. The prevalence threshold is study and sample type specific but usually varies between 0.01 (1% of the samples) and 0.1 (10% of the samples). 
The minimum ASV count usually varies between 1-10.
```{r}
# set parameters
prevalenceTheshold <- 0.02 # 2% of the samples
mincountsThreshold <- 1 # 1 ASV counts

# check ASVs counts
print(paste('before Rare ASVs removal: ',ntaxa(bact)))
# removing 
bact <- filter_taxa(bact, function(x) sum(x>mincountsThreshold) > (prevalenceTheshold*length(x)), TRUE)
# check ASVs counts
print(paste('after  Rare ASVs removal: ',ntaxa(bact)))
```

6) Samples filtering based on read counts (second round) At this step, samples with a low number
of reads will be removed. The minimum read count threshold can vary depending on the sample type,
the microbial diversity in the sample, whether we are dealing with 16S/ITS data and other parameters. Threshold typically ranges between 2’000 and 10’000 ASV counts.
```{r}
# minimum read threshold
minreadsTreshold <- 2000

# check number of samples
print(paste('before low number reads sample removal: ',nsamples(bact)))

# filtering base on the threshold
bact <- prune_samples(colSums(otu_table(bact)) > minreadsTreshold, bact)

# check number of samples
print(paste('after  low number reads sample removal: ',nsamples(bact)))
```

7) Remove positive controls
We still have some positive controls in the samples, for later analysis they can become outliers. So we will remove them right here.
```{r}
# check number of samples
print(paste('before positive control removal: ',nsamples(bact)))

# filtering base on the threshold
bact <- prune_samples(grepl('ZymoCtrl', sample_names(bact)) == F, bact)

# check number of samples
print(paste('after positive control removal: ',nsamples(bact)))
```

plot the prevalence after all filtering steps
```{r}
pdf(file=paste0('/Users/senmnes/DESKTOP_HD/Bernet_6th/pre-process_plot','/AF.Bact1Y_top20.pdf'), width=8, height=4, useDingbats=F, compress=F)
#
  PRA(bact,'Bacteria') + labs(title = '')
  
dev.off()
```

### Preview the sub groups
```{r}
print('all samples')
table(sample_data(bact)$Group)

# 1YEAR groups
bact.1Y <- prune_samples(sample_data(bact)$SampleCollection=='BCNasalswab1YEAR', bact)
print('collected at 1 Year')
table(sample_data(bact.1Y)$Group)

# ACUTE groups
bact.ACUTE <- prune_samples(sample_data(bact)$SampleCollection=='BCNasalswabAccute', bact)
print('collected at Acute')
table(sample_data(bact.ACUTE)$Group)
```


# Aim 1. 
## Alpha diversity analysis
Alpha diversity metrics summarize the structure of an ecological community with respect to its richness (number of taxonomic groups), evenness (distribution of abundances of the groups), or both. One commonly used measure is the Shannon index (taking into account both abundance and evenness).

Note: Alpha diversity is performed on non-normalized data.
```{r}
# init Shannon index calculation resuls
sample_data(bact.1Y)$Shannon <- estimate_richness(bact.1Y, split = TRUE, measures = c('Shannon'))[,1]
sample_data(bact.1Y)$Chao1 <- estimate_richness(bact.1Y, split = TRUE, measures = c('Chao1'))[,1]
# convert to dataframe for plot
bact1Y.data <- data.frame(sample_data(bact.1Y))
```

## Plot the diversity & richness
```{r}
aim1_path <- '/Users/senmnes/DESKTOP_HD/Bernet_6th/aim1_plots'
#### Diversity
png(file = paste0(aim1_path,'/Aim1_bact_Diversity.png'), width = 1210, height = 1210, unit='px')
# 
ggplot(bact1Y.data, aes(x = Group, y = Shannon)) +
    geom_boxplot(aes(y = Shannon,color=Group,fill=Group),width=0.7,outlier.size = -1) + 
    geom_jitter(aes(y = Shannon,fill = Group,alpha=0.95),
                shape = 21,width = 0.15,size = 6) +
    stat_compare_means(hjust=-0.25,vjust=0.5,size= 12)+
    labs(x='', y='Shannon index',
         title = paste0('Baterial diversity')) +
    scale_fill_manual(values = adjustcolor(custom_palette,alpha=0.65))  +
    scale_color_manual(values = adjustcolor(custom_palette)) + 
    ylim(0,max(bact1Y.data$Shannon * 1.14)) + theme_light()+ 
      theme(legend.position = 'right',
          legend.background = element_rect(linetype="solid",colour ="black"),
          text = element_text(size=45),legend.key.size = unit(2, "cm"))+
    guides(alpha="none")

dev.off()

#### Richness
png(file = paste0(aim1_path,'/Aim1_bact_Richness.png'), width = 1210, height = 1210, unit='px')
# 
print(ggplot(bact1Y.data, aes(x = Group, y = Chao1)) +
    geom_boxplot(aes(y = Chao1,color=Group,fill=Group),width=0.7,outlier.size = -1) + 
    geom_jitter(aes(y = Chao1,fill = Group,alpha=0.95),
                shape = 21,width = 0.15,size = 6) +
    stat_compare_means(hjust=-0.25,vjust=0.5,size= 12)+
    labs(x='', y='Chao1 index',
         title = paste0('Bacterial richness')) +
    scale_fill_manual(values = adjustcolor(custom_palette, alpha = 0.65))  +
    scale_color_manual(values = adjustcolor(custom_palette)) + 
    ylim(0,max(bact1Y.data$Chao1 * 1.14)) +theme_light()+ 
      theme(legend.position = 'right',
          legend.background = element_rect(linetype="solid",colour ="black"),
          text = element_text(size=45),legend.key.size = unit(2, "cm"))+
    guides(alpha="none"))
dev.off()
```


## Normalization
Normalization of ASV data is key to enable meaningful comparison of data from different measurements visualize amplicons data effectively. The preferred method is Cumulative Sum Scaling (CSS) implemented in the metagenomeSeq package, a method accounting for sparsity due to under-sampling.
```{r}
# Convert into a normal distribution [1YEAR] 
bact.norm.1Y <- bact.1Y
# normalized otu table
otu_table(bact.norm.1Y) <-otu_table(MRcounts(cumNorm(phyloseq_to_metagenomeSeq(bact.norm.1Y),p=cumNormStatFast(phyloseq_to_metagenomeSeq(bact.norm.1Y))),norm=TRUE, log=TRUE), taxa_are_rows=TRUE)

sample_data(bact.norm.1Y)$Group <- factor(sample_data(bact.norm.1Y)$Group,levels=c('Healthy','Wheeze'))
```

## Beta diversity analysis
Beta diversity is a measure of how similar (or dissimilar) the samples are between eachother, and is usually represented by a distance matrix which is then used to do Principal Coordinates Analysis (PCoA). The result of this is an ordination plot of multiple dimensions, where each sample is a point and the distance between the points represents the similarity of those samples (closer together = more similar). When a phylogenetic tree is available (e.g. when running the dada2 pipeline), the unifrac distance can be calculated (taking into accounts phylogeny). Another common distance is Bray Curtis (bray), which can be used if a phylogenetic tree is not provided.

Note: Beta diversity is usually calculated on normalized data.

1) Visualization
The distance matrix is used to do Principal Coordinates Analysis (PCoA).
```{r}
# Weighted Unifrac
bact_ord_Wunifrac <- ordinate(bact.norm.1Y, 'PCoA',"unifrac", weighted=TRUE)

# make plot
png(file = paste0(aim1_path,'/Aim1_Bacterial_Beta_Diversity.png'), width = 1410, height = 1410, unit = 'px')
#
print(plot_ordination(bact.norm.1Y, bact_ord_Wunifrac, color='Group') +
          stat_ellipse(geom = "polygon", linetype=0, alpha=0.25, aes(fill=Group))+
          geom_point(size = 7) +
          labs(title = paste0('Bacteria weighted Unifrac PCoA')) +
            xlab(paste0('PCoA1 ',round(bact_ord_Wunifrac$values$Relative_eig*100)[1],'%'))+
            ylab(paste0('PCoA2 ',round(bact_ord_Wunifrac$values$Relative_eig*100)[2],'%'))+
            scale_fill_manual(values = adjustcolor(custom_palette, alpha = 0.6))  +
            scale_color_manual(values = adjustcolor(custom_palette)) +
            geom_xsidedensity(aes(fill=Group),alpha=0.55) + 
            geom_ysidedensity(aes(fill=Group),alpha=0.55) + theme_light() +
            theme(axis.text.x=element_blank(),
                  axis.ticks.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks.y=element_blank()) +
            theme(legend.position = c(0.79, 0.1),legend.title = element_blank(),
                  legend.background = element_rect(linetype="solid",colour ="black"),
                  text = element_text(size=45),legend.key.size = unit(2, "cm")))
#
dev.off()
```

## Differential abundance testing with Maaslin2
```{r}
# build maasline analysis result
aim1_mas_bact <- Maaslin2(
    input_data = otu_table(bact.norm.1Y), 
    input_metadata = as.data.frame(sample_data(bact.norm.1Y)), 
    min_prevalence = 0.0,
    normalization = "NONE",
    transform = 'NONE',
    output = "aim1", 
    max_significance = 0.05,
    fixed_effects = c("Group"),
    standardize = FALSE)

# make a df for plot
aim1_mas_bact <- aim1_mas_bact$results
aim1_mas_bact <- aim1_mas_bact %>%
    dplyr::filter(qval <= 0.05) %>%
rownames_to_column(var = "id")
# add new variable
aim1_mas_bact$is.neg <- ifelse(aim1_mas_bact$coef>=0,'Wheeze','Healthy')

#### LM coeffecient plot
png(file = paste0(aim1_path,'/Aim1_bact_Differential_testing.png'), width = 1410, height = 880, unit='px')
#
print(ggplot(aim1_mas_bact, aes(x = reorder(feature,-coef), y = coef, fill = is.neg)) +
  geom_bar(stat="identity") +
  coord_flip() +
  geom_hline(yintercept = 0, linetype="dotted") + 
  ylim(-3, 3) +
  labs(x='', y='LM coefficient',title='Bacterial DA testing')+
  scale_fill_manual(values = adjustcolor(c("cadetblue4","lightsalmon3")),
                                         labels=c('Healthy','Wheeze')) + 
  theme_light() + 
  theme(legend.position = c(0.89,0.88),legend.title = element_blank(),
        legend.background = element_rect(linetype="solid",colour ="black"),
        text = element_text(size=30),legend.key.size = unit(2, "cm")))
#
dev.off()
```

## PSTA
plot significant taxa abundance
```{r}
PSTA <- function(taxa,factor,norm.plsq){
  df_holder <- c()
  # subset sample
  norm.plsq<-prune_samples(is.na(sample_data(norm.plsq)[[factor]])==F,norm.plsq)
  atr <- unique(sample_data(norm.plsq)[[factor]])
  if(length(atr) !=1){
    for(i in atr){
      i_sta <- prune_samples(sample_data(norm.plsq)[[factor]] == i,norm.plsq)
      # subset taxa
      i_sta <- prune_taxa(taxa_names(norm.plsq) == taxa, i_sta)
      # otu df
      i_otu<- as.data.frame(otu_table(i_sta))
      i_otu <- i_otu%>% pivot_longer(everything(),names_to='Sample',values_to = 'count')
      i_otu$type <- i
      # append to list
      df_holder[[length(df_holder)+1]] <- i_otu
    }
  }
  if(length(atr)==2){
    color <-c('cadetblue4','lightsalmon3')
  }else{
    color <-c('seagreen3','orange3','purple3','lightsteelblue4')
  }
  # merge df
  all_xy <- df_holder %>% reduce(full_join)
  
  # boxplot
  plt<-ggplot(all_xy, aes(type,count)) + 
    geom_boxplot(aes(y = count, fill = type), alpha=0.6,width= 0.7,outlier.size = -1)+
    geom_jitter(aes(fill = type),shape = 21,width = 0.15,size=7) +
    labs(color = 'Group',fill = 'Group',x='',y='Abundance(logCSS)', title = taxa) +
    scale_fill_manual(values=color) +
    theme_light() + theme(legend.position = 'right',
                          plot.title = element_text(face = "bold"),
                          legend.background = element_rect(linetype="solid",
                                                           colour ="black"),
                          text = element_text(size=30),legend.key.size = unit(2, "cm"))
  
  return(plt)
}
```

```{r}
png(file = paste0(aim1_path,'/bact_PSTA_12.png'),width = 1210, height = 1210, unit='px')
  PSTA('s.Moraxella_catarrhalis_12','Group',bact.norm.1Y) + stat_compare_means(paired = FALSE,hjust = -0.25, size=10)
dev.off()

png(file = paste0(aim1_path,'/bact_PSTA_45.png'),width = 1210, height = 1210, unit='px')
  PSTA('s.Gemella_haemolysans_45','Group',bact.norm.1Y)
dev.off()
```

# Aim 2.
## Pearson correlation test
- Numerical type records
```{r}
numerical.variables <-c('AbxCoursesPregnancy','GestAgeBirth','WeightBirth','ChestInfections1YEAR','Colds1YEAR','DiagnosedBronchiolitis1YEAR','AbxSinceBorn1YEAR','NbHouse')
```

```{r}
pearson.corr_test<-function(plsq_df){
  # result df
  res_df <- data.frame('factor'=numerical.variables,'correlation'=rep(0,length(numerical.variables)),'pearson.p.value' = rep(0,length(numerical.variables)))
  #
  for(f in numerical.variables){
    plsq_df[[f]] <- as.numeric(plsq_df[[f]])
    variable <- unique(plsq_df[[f]])
      if(length(variable)!=0){
        # Corr test
        res <- cor.test(plsq_df$Shannon, plsq_df[[f]])
        # append value
        res_df[(res_df$factor == f),]$pearson.p.value <- res$p.value
        res_df[(res_df$factor == f),]$correlation <- res$estimate[[1]]
      }else{
        res_df[(res_df$factor == f),]$wilcox.p.value <- NA
        res_df[(res_df$factor == f),]$correlation <- NA
      }
  }
  return(res_df)
}

Bact.pc <- pearson.corr_test(bact1Y.data)
```

## Wilcoxon test
```{r}
wilconxon_test<-function(plsq_df){
  factor_sets <- c(genetic_factors,birth_pregnancy_factors,clinical_factors,environmental_factors)
  # result df
  res_df <- data.frame('factor'=factor_sets,'wilcox.p.value'=rep(0,length(factor_sets)),'W' = rep(0,length(factor_sets)))
  #
  for(f in factor_sets){
    plsq_df[[f]] <- as.factor(plsq_df[[f]])
    
    if(f %in% numerical.variables == F){
      variable <- unique(plsq_df[[f]])
      if(length(variable)>=2){
        # subset Shannon index by variable
        v1_shannon <- plsq_df[(plsq_df[[f]] == variable[1]),]$Shannon
        v2_shannon <- plsq_df[(plsq_df[[f]] == variable[2]),]$Shannon
        # Wilconxon test
        res <- wilcox.test(v1_shannon, v2_shannon)
        # append value
        res_df[(res_df$factor == f),]$wilcox.p.value <- res$p.value
        res_df[(res_df$factor == f),]$W <- res$statistic[[1]]
      }else{}
    }else{
      res_df[(res_df$factor == f),]$wilcox.p.value <- NA
      res_df[(res_df$factor == f),]$W <- NA
    }
  }
  return(res_df)
}
#
Bact.wt<-wilconxon_test(bact1Y.data)
```

## Alpha Diversity
### HospitalAdmissionsAsthmaWheeze1YEAR
```{r}
aim2_path <- '/Users/senmnes/DESKTOP_HD/Bernet_6th/aim2_plot'
#### Diversity
# remove any NAs
HAAW.data <- bact1Y.data[is.na(bact1Y.data$HospitalAdmissionsAsthmaWheeze1YEAR) == F,]

# pdf(file = paste0(aim2_path,'/Aim2_AD_HAAW.pdf'), width = 5.5, height = 5, useDingbats = F, compress = F)
# #
# ggplot(HAAW.data, aes(x = HospitalAdmissionsAsthmaWheeze1YEAR, y = Shannon)) +
#     geom_boxplot(aes(y = Shannon),width=0.7,outlier.size = -1) + 
#     geom_jitter(aes(y = Shannon,fill = Group,alpha=0.95),shape = 21,width = 0.15) +
#     stat_compare_means(comparisons = list(unique(HAAW.data$HospitalAdmissionsAsthmaWheeze1YEAR)),paired = FALSE,vjust = -0.5)+
#     labs(x='', y='Shannon index',
#          title = paste0('Bacterial diversity'),
#          subtitle = ('Hospital visits for asthma or wheezing')) +
#     scale_fill_manual(values = adjustcolor(custom_palette,alpha=0.65))  +
#     scale_color_manual(values = adjustcolor(custom_palette)) + 
#     ylim(0,max(HAAW.data$Shannon * 1.14)) +theme_light()+
#   theme(legend.position = 'right',legend.title = element_blank(),
#                   legend.background = element_rect(linetype="solid",colour ="black")) +
#   guides(alpha="none")
# #
# dev.off()
```

### Breastfeeding Birth
```{r}
#### Diversity
# remove any NAs
BFB.data <- bact1Y.data[is.na(bact1Y.data$BreastfeedingBirth) == F,]

png(file = paste0(aim2_path,'/Aim2_AD_BFB.png'), width = 1210, height = 1210, unit='px')
#
ggplot(BFB.data, aes(x = BreastfeedingBirth, y = Shannon)) +
    geom_boxplot(aes(y = Shannon),width=0.7,outlier.size = -1) + 
    geom_jitter(aes(y = Shannon,fill = Group),shape = 21,
                width = 0.15,size=7) +
    stat_compare_means(paired = FALSE,vjust = 0.5, hjust = -0.25, size = 13)+
    labs(x='', y='Shannon index',
         title = paste0('Bacterial diversity'),
         subtitle = ('Breastfeeding after birth')) +
    scale_fill_manual(values = adjustcolor(custom_palette))  +
    scale_color_manual(values = adjustcolor(custom_palette)) +
    ylim(0,max(BFB.data$Shannon * 1.14)) + 
    theme_bw() + theme(legend.position = 'right',
                       legend.background = element_rect(linetype="solid",colour ="black"),
                          plot.subtitle = element_text(face = "bold",hjust=0.5),
                          text = element_text(size=30),legend.key.size = unit(2, "cm"))+
  guides(alpha="none")
#
dev.off()
```

### Chest infections
```{r}
#### Diversity
png(file = paste0(aim2_path,'/Aim2_BD_Chestinfection.png'), width = 1210, height = 1210, unit='px')
# 
ggplot(bact1Y.data, aes(x = ChestInfections1YEAR, y = Shannon)) +
    geom_smooth(method = lm, span = 0.1, size = 5)+
    geom_point(aes(y = Shannon,fill = Group),shape = 21,width = 0.15,size=7) +
    stat_cor(method = "pearson",size = 13,hjust=-0.8) +
    labs(x='', y='Shannon index',
         title = 'Bacterial diversity',
         subtitle = paste0('Number of chest infections before 1YEAR')) +
    scale_fill_manual(values = adjustcolor(custom_palette))  +
    scale_color_manual(values = adjustcolor(custom_palette)) + 
    ylim(0,max(bact1Y.data$Shannon * 1.14)) +theme_light()+
    theme_bw() + theme(legend.position = 'right',
                       legend.background = element_rect(linetype="solid",colour ="black"),
                          plot.subtitle = element_text(face = "bold",hjust=0.5),
                          text = element_text(size=30),legend.key.size = unit(2, "cm"))+
  guides(alpha="none")
#
dev.off()
```

### Colds1YEAR
```{r}
#### Diversity
png(file = paste0(aim2_path,'/Aim2_BD_Colds1YEAR.png'), width = 1210, height = 1210, units='px')
# 
ggplot(bact1Y.data, aes(x = Colds1YEAR, y = Shannon)) +
    geom_smooth(method = lm, span = 0.1, size = 5)+
    geom_point(aes(y = Shannon,fill = Group),shape = 21,width = 0.15,size=7) +
    stat_cor(method = "pearson",size=13, hjust=-0.8) +
    labs(x='', y='Shannon index',
         title = 'Bacterial diversity',
         subtitle = paste0('Number of colds before 1YEAR')) +
    scale_fill_manual(values = adjustcolor(custom_palette))  +
    scale_color_manual(values = adjustcolor(custom_palette)) + 
    ylim(0,max(bact1Y.data$Shannon * 1.14))+
    theme_bw() + theme(legend.position = 'right',
                       legend.background = element_rect(linetype="solid",colour ="black"),
                          plot.subtitle = element_text(face = "bold",hjust=0.5),
                          text = element_text(size=30),legend.key.size = unit(2, "cm"))+
  guides(alpha="none")
#
dev.off()
```

### AbxSinceBorn1YEAR
```{r}
#### Diversity
png(file = paste0(aim2_path,'/Aim2_BD_AbxSinceBorn1YEAR.png'), width = 1210, height = 1210, units = 'px')
# 
ggplot(bact1Y.data, aes(x = AbxSinceBorn1YEAR, y = Shannon)) +
    geom_smooth(method = lm, span = 0.1, size = 5)+
    geom_point(aes(y = Shannon,fill = Group),shape = 21,width = 0.15,size=7) +
    stat_cor(method = "pearson",size=13,hjust=-0.8) +
    labs(x='', y='Shannon index',
         title = 'Bacterial diversity',
         subtitle = paste0('Number of antibiotics used before 1YEAR')) +
    scale_fill_manual(values = adjustcolor(custom_palette))  +
    scale_color_manual(values = adjustcolor(custom_palette)) + 
    ylim(0,max(bact1Y.data$Shannon * 1.14)) +
    theme_bw() + theme(legend.position = 'right',
                       legend.background = element_rect(linetype="solid",colour="black"),
                          plot.subtitle = element_text(face = "bold",hjust=0.5),
                          text = element_text(size=30),legend.key.size = unit(2, "cm"))+
  guides(alpha="none")
#
dev.off()
```


## Aim2 ANOSIM
```{r}
anosim_test <- function(list,plsq_obj,form){
  set.seed(2)
  # fixed setting
  distance <- 'wunifrac'
  # test all genetic factors
  design <- paste(list, collapse = " + ")
  # build formula
  an_data <- as.data.frame(as.matrix(sample_data(plsq_obj)))
  # generate the model
  test <- adonis2(formula = as.formula(paste(form, design)),data = as.data.frame(as.matrix(sample_data(plsq_obj))), distance = 'wunifrac', dfun = vegdist,na.action=na.omit)
  test <- as.data.frame(test)
  # sort by Pr(>F)
  test <- test[order(test$`Pr(>F)`),]
  print(test)
}

# 1YEAR groups
# genetic factor
form <- 'phyloseq::distance(bact.norm.1Y, method=distance) ~'
g.f.anosim <- anosim_test(genetic_factors,bact.norm.1Y,form)
# brith factor
b.g.anosim <- anosim_test(birth_pregnancy_factors,bact.norm.1Y,form)
# clinical factor
c.f.anosim <- anosim_test(clinical_factors,bact.norm.1Y,form)
# environmental factor
e.f.anosim <- anosim_test(environmental_factors,bact.norm.1Y,form)
```

## Aim2 DA testing
DA testing feature number counter
```{r}
FNC <- function(factor_list,norm.plsq){
  set.seed(2)
  # init df
  holder_df <- data.frame('factor' = c(factor_list), 'counts' = rep(0,length(factor_list)))
  #
  for(i in factor_list){
   aim_mas<-Maaslin2(
        input_data = otu_table(norm.plsq), 
        input_metadata = as.data.frame(sample_data(norm.plsq)), 
        min_prevalence = 0.0,
        normalization = "NONE",
        transform = 'NONE',
        output = "FNC", 
        max_significance = 0.05,
        fixed_effects = c(i),
        standardize = FALSE,
        plot_heatmap = TRUE)
   # make a df
    aim_mas <- aim_mas$results
    aim_mas <- aim_mas %>%
        dplyr::filter(qval < 0.05) %>%
    rownames_to_column(var = "id")
    #
    holder_df[(holder_df$factor == i),]$counts <- nrow(aim_mas)
  }
  return(holder_df)
}
```

- Outputs the feature counts result
```{r}
bact.genetics <- FNC(genetic_factors,bact.norm.1Y)
bact.brith_n_pre <- FNC(birth_pregnancy_factors,bact.norm.1Y)
#
bact.clinical <- FNC(clinical_factors,bact.norm.1Y)
bact.environment <- FNC(environmental_factors,bact.norm.1Y)
```

### NbHouse
```{r}
# a2_1Y_mas <- Maaslin2(
#     input_data = otu_table(bact.norm.1Y), 
#     input_metadata = as.data.frame(sample_data(bact.norm.1Y)), 
#     min_prevalence = 0.0,
#     normalization = "NONE",
#     transform = 'NONE',
#     output = "aim2.Nbhouse", 
#     max_significance = 0.05,
#     fixed_effects = c('NbHouse'),
#     standardize = FALSE)
# 
# # make a df
# a2_1Y_mas <- a2_1Y_mas$results
# a2_1Y_mas <- a2_1Y_mas %>%
#     dplyr::filter(qval < 0.05) %>%
# rownames_to_column(var = "id")
# # add new variable
# a2_1Y_mas$is.neg <- ifelse(a2_1Y_mas$coef>=0,'Wheeze','Healthy')
# 
# # Colds1YEAR
# pdf(file = paste0(aim2_path,'/Bacteria_Aim2_DA_NBhouse.pdf'), width = 5, height = 5, useDingbats = F, compress = F)
# #
# ggplot(a2_1Y_mas, aes(x = reorder(feature,-coef), y = coef, fill = is.neg)) +
#   geom_bar(stat="identity") +
#   coord_flip() +
#   geom_hline(yintercept = 0, linetype="dotted") + 
#   ylim(min(a2_1Y_mas$coef)*1.1, max(a2_1Y_mas$coef)*1.1) +
#   labs(x='Taxonomy', y='LM coefficient',
#          title = ('Bacterial'),
#          subtitle = paste0('Number of people living in the house'))+
#   scale_fill_manual(values = adjustcolor(c('deeppink4','pink2'),alpha=0.5))+ 
#   theme_light() + theme(legend.position="none")
# #
# dev.off()
```

### DayCare1YEAR
```{r}
a2_1Y_mas <- Maaslin2(
    input_data = otu_table(bact.norm.1Y), 
    input_metadata = as.data.frame(sample_data(bact.norm.1Y)), 
    min_prevalence = 0.0,
    normalization = "NONE",
    transform = 'NONE',
    output = "aim2.DayCares", 
    max_significance = 0.05,
    fixed_effects = c('DayCare1YEAR'),
    standardize = FALSE,
    plot_heatmap = TRUE)

# make a df
a2_1Y_mas <- a2_1Y_mas$results
a2_1Y_mas <- a2_1Y_mas %>%
    dplyr::filter(qval < 0.05) %>%
rownames_to_column(var = "id")
# add new variable
a2_1Y_mas$is.neg <- ifelse(a2_1Y_mas$coef>=0,'Daycare','Absent')

# DayCare1YEAR
png(file = paste0(aim2_path,'/Bacteria_Aim2_DA_DayCare1YEAR.png'), width = 1410, height = 880, units='px')
#
ggplot(a2_1Y_mas, aes(x = reorder(feature,-coef), y = coef, fill = is.neg)) +
  geom_bar(stat="identity") +
  coord_flip() +
  geom_hline(yintercept = 0, linetype="dotted") + 
  ylim(min(a2_1Y_mas$coef)*1.1, max(a2_1Y_mas$coef)*1.1) +
  labs(x='', y='LM coefficient',
         title = ('Bacteria DA testing'),
         subtitle = paste0('Attending daycare before 1YEAR'))+
  scale_fill_manual(values = adjustcolor(c('lightsteelblue2','lightsteelblue4')),
                                         labels=c('Absent','Daycare')) + 
  theme_bw() + 
  theme(legend.position = c(0.89,0.88),legend.title = element_blank(),
        legend.background = element_rect(linetype="solid",colour ="black"),
        plot.subtitle = element_text(face = "bold",hjust=0.5),
        text = element_text(size=30),legend.key.size = unit(2, "cm"))
#
dev.off()
```

### TimeOfYear
```{r}
# a2_1Y_mas <- Maaslin2(
#     input_data = otu_table(bact.norm.1Y), 
#     input_metadata = as.data.frame(sample_data(bact.norm.1Y)), 
#     min_prevalence = 0.0,
#     normalization = "NONE",
#     transform = 'NONE',
#     output = "aim2.TimeOfYear", 
#     max_significance = 0.05,
#     fixed_effects = c('TimeOfYear'),
#     standardize = FALSE,
#     plot_heatmap = TRUE)
# 
# # make a df
# a2_1Y_mas <- a2_1Y_mas$results
# a2_1Y_mas <- a2_1Y_mas %>%
#     dplyr::filter(qval < 0.05) %>%
# rownames_to_column(var = "id")
# # add new variable
# a2_1Y_mas$is.neg <- ifelse(a2_1Y_mas$coef>=0,'Wheeze','Healthy')
# 
# # DayCare1YEAR
# pdf(file = paste0(aim2_path,'/Bacteria_Aim2_DA_TimeOfYear.pdf'), width = 7, height = 5, useDingbats = F, compress = F)
# #
# ggplot(a2_1Y_mas, aes(x = reorder(feature,-coef), y = coef, fill = is.neg)) +
#   geom_bar(stat="identity") +
#   coord_flip() +
#   geom_hline(yintercept = 0, linetype="dotted") + 
#   ylim(min(a2_1Y_mas$coef)*1.1, max(a2_1Y_mas$coef)*1.1) +
#   labs(x='', y='LM coefficient',
#          title = ('Bacterial'),
#          subtitle = paste0('Temperature of the sampling season'))+
#   scale_fill_manual(values = adjustcolor(c('lightsteelblue2','lightsteelblue4')),
#                                          labels=c('Healthy','Wheeze')) + 
#   theme_light() + theme(legend.position = c(0.88, 0.90),legend.title = element_blank(),
#                   legend.background = element_rect(linetype="solid",colour ="black"))
# #
# dev.off()
```

# Aim 3, Characterise the bact community development with children’s Age(time model). 
Get the patient has all sample across the collection time
```{r}
aim3_team <- c('6045EC','6053LLD','6078MMH','4087AT','4086AMR','4117KF','6113ML','6133HC','7242MJH','7276HC','1317DR','6443LSR','4461SM','6469AB','1494OEM','7518GG','7527RM','5533LB','4565ES','7626RCH','6661LD','1664NC','5681KP','5693FN','6701JA','4710VN','4776AK','7784SLH','5788RA','5804SA','7845DTA')

team_012MTH <- long_ss[(long_ss$SampleCollection != 'Acute'|long_ss$SampleCollection != 'Acute-NP'),]$Barcode
aim3_plsq <- prune_samples(sample_data(bact)$Barcode %in% team_012MTH, bact)
aim3_plsq <- prune_samples(sample_data(aim3_plsq)$Patient %in% aim3_team, aim3_plsq)

# calculate Chao1 & Shannon index
sample_data(aim3_plsq)$Chao1 <- estimate_richness(aim3_plsq, split = TRUE, measures = c('Chao1'))[,1]
sample_data(aim3_plsq)$Shannon <- estimate_richness(aim3_plsq, split = TRUE, measures = c('Shannon'))[,1]

# real Age, Age for each sample
sample_data(aim3_plsq)$AgeByMonth <- 0
for(i in seq_along(sample_data(aim3_plsq)$Patient)){
  bd <- sample_data(aim3_plsq)$ParticipantsDateofbirth[i]
  if(sample_data(aim3_plsq)$SampleCollection[i] == 'BCNasalswabBaseline'){
    sd <- sample_data(aim3_plsq)$DateofVisitBaseline[i]
    #print(bd)
    sample_data(aim3_plsq)$AgeByMonth[i] <- day(as.period(interval(start=bd,end=sd), unit='days'))/30.437
  }else if(sample_data(aim3_plsq)$SampleCollection[i] == 'BCNasalswab3MTH'){
    sd <- sample_data(aim3_plsq)$DateofVisit3MTH[i]
    sample_data(aim3_plsq)$AgeByMonth[i] <- day(as.period(interval(start=bd,end=sd), unit='days'))/30.437
  }else if(sample_data(aim3_plsq)$SampleCollection[i] == 'BCNasalswab6MTH'){
    sd <- sample_data(aim3_plsq)$DateofVisit6MTH[i]
    sample_data(aim3_plsq)$AgeByMonth[i] <- day(as.period(interval(start=bd,end=sd), unit='days'))/30.437
  }else if(sample_data(aim3_plsq)$SampleCollection[i] == 'BCNasalswab1YEAR'){
    sd <- sample_data(aim3_plsq)$DateofVisit1YEAR[i]
    sample_data(aim3_plsq)$AgeByMonth[i] <- day(as.period(interval(start=bd,end=sd), unit='days'))/30.437
  }
}

sample_data(aim3_plsq)$AgeGroup <- NA
# set attribute for Age groups
for(k in seq_along(sample_data(aim3_plsq)$AgeByMonth)){
  if((sample_data(aim3_plsq)$SampleCollection[k]=='BCNasalswabBaseline'))
    {sample_data(aim3_plsq)$AgeGroup[k] <- 'T1-1week'}
  else if((sample_data(aim3_plsq)$SampleCollection[k]=='BCNasalswab3MTH'))
    {sample_data(aim3_plsq)$AgeGroup[k] <- 'T2-3months'}
  else if((sample_data(aim3_plsq)$SampleCollection[k]=='BCNasalswab6MTH'))
    {sample_data(aim3_plsq)$AgeGroup[k] <- 'T3-6months'}
  else if((sample_data(aim3_plsq)$SampleCollection[k]=='BCNasalswab1YEAR'))
    {sample_data(aim3_plsq)$AgeGroup[k] <- 'T4-1year'}
}

# convert into factor value
sample_data(aim3_plsq)$AgeGroup <- factor(sample_data(aim3_plsq)$AgeGroup, levels=c('T1-1week','T2-3months','T3-6months','T4-1year'))
```

## Alpha diversity analysis
```{r}
aim3_palette <- c('seagreen3','orange3','purple3','lightsteelblue4')
aim3_path <- '/Users/senmnes/DESKTOP_HD/Bernet_6th/aim3_plot'
#### Diversity
pdf(file = paste0(aim3_path,'/Aim3_BD_AgeGroup.pdf'), width = 5, height = 5, useDingbats = F, compress = F)
#
ggplot(sample_data(aim3_plsq),aes(x=Group, y=Shannon)) + 
        geom_boxplot(aes(y = Shannon)) +
        geom_jitter(aes(y = Shannon,fill = Group,alpha=0.95),shape = 21,width = 0.15) +
        stat_compare_means(comparisons = list(unique(sample_data(aim3_plsq)$Group)),
                           paired = FALSE,vjust = -0.5)+ facet_wrap(vars(AgeGroup))+
        labs(x='', y='Shannon index',
             title = 'Bacterial diversity',
             subtitle = paste0('Longitudinal group')) +
        scale_fill_manual(values = adjustcolor(custom_palette)) +
        scale_color_manual(values = adjustcolor(custom_palette)) +
        ylim(0,max(sample_data(aim3_plsq)$Shannon * 1.3))  +  theme_light() +
        theme(legend.position = 'right',
        legend.background = element_rect(linetype="solid",colour ="black")) +
        guides(alpha="none")
#
dev.off()

#### Richness
pdf(file = paste0(aim3_path,'/Aim3_BR_AgeGroup.pdf'), width = 5, height = 5, useDingbats = F, compress = F)
#
ggplot(sample_data(aim3_plsq),aes(x=Group, y=Chao1)) + 
        geom_boxplot(aes(y = Chao1)) +
        geom_jitter(aes(y = Chao1,fill = Group,alpha=0.95),shape = 21,width = 0.15) +
        stat_compare_means(comparisons = list(unique(sample_data(aim3_plsq)$Group)),
                           paired = FALSE,vjust = -0.5)+ facet_wrap(vars(AgeGroup))+
        labs(x='', y='Chao1 index',
             title = 'Bacterial richness',
             subtitle = paste0('Longitudinal group')) +
                scale_fill_manual(values = adjustcolor(custom_palette)) +
        scale_color_manual(values = adjustcolor(custom_palette)) +
        ylim(0,max(sample_data(aim3_plsq)$Chao1 * 1.3))  +  theme_light() +
        theme(legend.position = 'right',
        legend.background = element_rect(linetype="solid",colour ="black")) +
        guides(alpha="none")
#
dev.off()
```

## Beta diversity analysis
The new phyloseq is directly assign from `bact`, which has not been normalized.
```{r}
otu_table(aim3_plsq) <-otu_table(MRcounts(cumNorm(phyloseq_to_metagenomeSeq(aim3_plsq),p=cumNormStatFast(phyloseq_to_metagenomeSeq(aim3_plsq))),norm=TRUE, log=TRUE), taxa_are_rows=TRUE)
```

```{r}
aim3_palette <- c('seagreen3','orange3','purple3','lightsteelblue4')
aim3_path <- '/Users/senmnes/DESKTOP_HD/Bernet_6th/aim3_plot'
# get df
aim3_bact_ord_Wunifrac <- ordinate(aim3_plsq,'PCoA',"unifrac", weighted=TRUE)
#
aim3_pcoA_DF <- plot_ordination(aim3_plsq, aim3_bact_ord_Wunifrac, color='AgeGroup',justDF = T)
# plotting
png(file = paste0(aim3_path,'/Aim1_bact_Beta_Diversity.png'),width=1410, height=1410, units='px')
#
ggplot(aim3_pcoA_DF,aes(Axis.1, Axis.2)) + 
  geom_point(aes(fill=AgeGroup,shape = Group), width = 0.15,size=7) +
  stat_ellipse(aes(fill = AgeGroup),geom = "polygon", linetype=0, alpha=0.15)+
  labs(color='TimeFactors', fill ='TimeFactors',
                 title = paste0('Bacteria weighted Unifrac PCoA'),
                 subtitle = 'Longitudinal group') +
              xlab(paste0('PCoA1 ',round(aim3_bact_ord_Wunifrac$values$Relative_eig*100)[1],'%'))+
              ylab(paste0('PCoA2 ',round(aim3_bact_ord_Wunifrac$values$Relative_eig*100)[2],'%'))+
              scale_fill_manual(values = adjustcolor(aim3_palette)) +
              scale_color_manual(values = adjustcolor(aim3_palette)) +
              scale_shape_manual(values=c(21,23)) + 
              geom_xsidedensity(aes(fill=AgeGroup),alpha=0.45) + 
              geom_ysidedensity(aes(fill=AgeGroup),alpha=0.45) +
              guides(size='none') + theme_light() +
              theme(axis.text.x=element_blank(),
                  axis.ticks.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks.y=element_blank())+
              theme(legend.position = 'right',legend.title = element_blank(),
                     legend.background = element_rect(linetype="solid",colour ="black"),
                  text = element_text(size=45),legend.key.size = unit(2, "cm"))+
        guides(alpha="none")
# 
dev.off()
```

## Aim3 DA testing
### AgeGroup
```{r}
a3_1Y_mas <- Maaslin2(
    input_data = otu_table(aim3_plsq), 
    input_metadata = as.data.frame(sample_data(aim3_plsq)), 
    min_prevalence = 0.0,
    normalization = "NONE",
    transform = 'NONE',
    output = "aim3.AgeGroup", 
    max_significance = 0.05,
    fixed_effects = c('AgeGroup'),
    standardize = FALSE)

# make a df
a3_1Y_mas <- a3_1Y_mas$results
a3_1Y_mas <- a3_1Y_mas %>%
    dplyr::filter(qval < 0.05) %>%
rownames_to_column(var = "id")
# add new variable
a3_1Y_mas$is.neg <- ifelse(a3_1Y_mas$coef>=0,'Wheeze','Healthy')

# AgeGroup
png(file = paste0(aim3_path,'/Bacteria_Aim3_DA_AgeGroup.png'), width = 1210, height = 880, units = 'px')
#
ggplot(a3_1Y_mas, aes(x = reorder(feature,-coef), y = coef, fill = is.neg)) +
  geom_bar(stat="identity") +
  coord_flip() +
  geom_hline(yintercept = 0, linetype="dotted") + 
  #ylim(min(a3_1Y_mas$coef)*1.1, max(a3_1Y_mas$coef)*1.1) +
  labs(x='Taxonomy', y='LM coefficient',
         title = ('Bacteria DA testing'),
         subtitle = paste0('AgeGroup'))+
  scale_fill_manual(values =c('pink2','pink4'))+ 
  theme_bw() + 
  theme(legend.position = 'none',legend.title = element_blank(),
        legend.background = element_rect(linetype="solid",colour ="black"),
        text = element_text(size=30),legend.key.size = unit(2, "cm"))
#
dev.off()
```

### Plot relative abundance of significant taxa
```{r}
png(file = paste0(aim3_path,'/bact_Age_Group_PSTA_14.png'), width = 1210, height = 1210, units='px')
  PSTA('g.Staphylococcus_14','AgeGroup',aim3_plsq)
dev.off()

png(file = paste0(aim3_path,'/bact_Age_Group_PSTA_58.png'), width = 1210, height = 1210, units='px')
  PSTA('s.Staphylococcus_epidermidis_58','AgeGroup',aim3_plsq)
dev.off()

png(file = paste0(aim3_path,'/bact_Age_Group_PSTA_25.png'), width = 1210, height = 1210, units='px')
  PSTA('s.Dolosigranulum_pigrum_25','AgeGroup',aim3_plsq)
dev.off()

png(file = paste0(aim3_path,'/bact_Age_Group_PSTA_12.png'), width = 1210, height = 1210, units='px')
  PSTA('s.Moraxella_catarrhalis_12','AgeGroup',aim3_plsq)
dev.off()
```


# Output Phyloseq Object to local
```{r}
#saveRDS(bact, '/Users/senmnes/DESKTOP_HD/Bernet_6th/bact.rds')
#saveRDS(aim3_plsq, '/Users/senmnes/DESKTOP_HD/Bernet_6th/aim3.bact.rds')
```